\documentclass[../main.tex]{subfiles}
\graphicspath{{\subfix{../}}}
\begin{document}

\chapter{Changelog}

The following changes have been made from draft version 2.0 to address
the comments received during the review process:


\begin{itemize}
\item \autoref{ch:Chapter 3} in this version covers what was Chapter 3 in the previous and includes also the problem statement and the material in the previous chapter on IMF ecosystem.
  \item \autoref{ch:Chapter 2}, previously Chapter 2, which includes background information on system, aspect, and information model, has been rewritten 
        and extended.
        \item \autoref{ch:IMF Language: Overview} has been updated to reflect new developments.
  \item \autoref{ch:How to Create an IMF Model}, which describes how to model, has been updated.

        \begin{itemize}
          \item Updated to reflect comments received from external draft review.
          \item Updated to reflect IMF changes as specified in \autoref{ch:The IMF Language Formalized}.
        \end{itemize}
  \item \autoref{ch:How to Specify IMF Types}, which describes how to create IMF types has been updated.

        \begin{itemize}
          \item Updated to reflect comments received from external draft review.
          \item Updated to reflect IMF changes as specified in \autoref{ch:The IMF Language Formalized}.
        \end{itemize}
        \item \autoref{ch:The IMF Language Formalized}, which includes the formal specification of the IMF language, has been restructured.
  The IMF ontology has been updated according to best practices for ontology engineering, including textual definitions for all central terms. 
  
  
        
        
        \end{itemize}
\end{document}
      